---
title: "visNetwork_tutorial"
author: "Nic Fox"
date: "01/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Visualise how nodes are linked with this interactive network visualisation using the `visNetwork` R library

Based on this tutorial: https://rpubs.com/a_pear_9/visNetwork

## Import the visNetwork library
```{r import_packages}
library(visNetwork)
```

## Create nodes and edges
First, create a list of numbers from 1 to 16 to use as node IDs.
```{r create_nodes, message=FALSE}
# Create sixteen nodes
node_IDs <- 1:16

# View the sixteen nodes
node_IDs
```

Then create a list of 'from' and 'to' node IDs to form the edges.
```{r create_edges, message=FALSE}
# create a list of numbers by taking each number in the 'population' variable
# twice (i.e. each number is repeated once) to give 32 numbers
from_nodes  <- rep(node_IDs, each = 2) # from

# create a list of numbers by taking a sample of 32 from the 'population' variable
# with replacement
to_nodes  <- sample(x=node_IDs, size=32, replace=TRUE) # to

# View the 'edge from' and 'to' values
cat('from_nodes: \n', from_nodes, '\n\n', 'to_nodes: \n', to_nodes)
```

Now, create a 'nodes' dataframe containing 2 columns:
- id (populated with the node IDs)
- label (populated with 'N' plus the node ID e.g. N1. This is what will be displayed in the node in the visualisation.)

Write the nodes dataframe to a CSV and view the dataframe.
```{r create_nodes_df, message=FALSE}
# Create dataframe of nodes
nodes <- data.frame(
                    id = node_IDs,
                    label = paste0("N", node_IDs),
                    title = paste0('This is node ', node_IDs)
)

# save the edges dataframe as a csv
write.csv(nodes, 'nodes.csv', row.names=FALSE)

# View the dataframe of nodes
nodes
```

Create an 'edges' dataframe with 2 columns:
- from (populated with the 'relate_fr' node IDs for the edges)
- to (populated with the 'relate_to' node IDs for the edges)

Then add a new 'title' columns containing the html text to be displayed in the tooltip (pop-up) when the user clicks on the edge.

Write the edges dataframe to a CSV and view the dataframe.
```{r create_edges_df, message=FALSE}
# Create dataframe of edges
edges <- data.frame(
  from = from_nodes,
  to   = to_nodes
)

# add a 'title' column contain the tooltip text for the edge
edges$title <- paste0('from: ', edges$from, '<br>to: ', edges$to)

# save the edges dataframe as a csv
write.csv(edges, 'edges.csv', row.names=FALSE)

# View the dataframe
edges
```

## Create and visualise the network
create a `visNetwork` using the nodes and edges dataframes and then define the aesthetics of the nodes and edges.

Then save the network visualisation as an html file, then view the visualisation.

Click on an edge to see the tooltip for the edge, and click on a node to see the tooltip for the node. If the tooltip doesn't appear on the first click, try once or twice more.

You can also zoom in and zoom out on the visualisation, and move nodes around.
```{r visualise_network, message=FALSE}
network <- visNetwork(nodes, edges) %>% 
                        visNodes(shape = "box", color = "#ffc34d", font = c(color="white")) %>% 
                        visEdges(color = "black")

# save network as an html file
visSave(network, file = "network.html", background = "white")

# view network
network
```